State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-10 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
fixed some typos in paper

Similar considerations need to be made regarding the representation of the dynamic field $v_t(\mathbf{r})$ using the basis function decomposition. The minimum distance between adjacent basis functions must also satisfy Shannon's sampling theorem to capture the dominant spectral information of the field. Thus, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_b < \frac{1}{2\rho_b\nu_{co}}. 
\end{equation}
where $\rho_b$ is an over-sampling parameter to determine the basis function separation and $\nu_{co}$ is the spatial cut-off frequency from the observations. For the Gaussian basis functions, $\phi(\mathbf{r})$, being considered, the basis function width parameter, $\sigma_{\phi}^2$, can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian is another Gaussian given by
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\nu}$, should be chosen to give at least 3 dB attenuation at $\boldsymbol\nu_{co}$ giving
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu}=\frac{2\boldsymbol\nu_c^\top \boldsymbol\nu_c}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_{co} $. Proofs for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.
% Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with at least the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma_{\phi}^2$.

The dynamic field must be spatially oversampled using a larger number of sensors then imposed by frequency analysis considerations, to use frequency analysis to determine the minimum distance and width of the basis functions. Following this, the distance between basis functions is usually larger then the distance between sensors, although this is not a requirement for using this estimation framework. \omg{Ken: can you please write something about what should be done in this case?} For a spatially homogeneous isotropic fields, the basis functions can be placed on a regular grid. Following this, the knowledge of the distance between basis functions directly implies the total number of basis functions required.

@@

	\lambda=\alpha^2( L+\kappa)- L 

@@

where $\alpha$, $\beta$ and $\kappa$ are constants (see ~\cite{Haykin2001} for more details). The sigma vectors are propagated through the system equations and weighted to form the predicted mean and covariance. The weights are calculated by 

@@

Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are  used to correct the predicted states. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the predicted cross-covariance matrix of the states is also calculated. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that as the disturbance, $e_t\left(\mathbf{r}\right)$ and the measurement noise $ \boldsymbol{\varepsilon}_t$ are additive the additive form of the URTS can be used , otherwise the state vector should be augmented with the noise terms. Using the additive form reduces the dimension and the number of sigma points used in the algorithm.

@@

\begin{figure*}[!th]
\centering 
\subfigure[][]{\includegraphics[width=2in]{./Graph/TrueField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimatedField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimationError90.pdf}}\\
\caption{One snapshot of the spatial field; a: True field; b: Estimated field; c: The absolute error.}
\label{fig:FieldEstimate}
\end{figure*}

>>>>

Similar considerations need to be made regarding the representation of the dynamic field $v_t(\mathbf{r})$ using the basis function decomposition. The minimum distance between adjacent basis functions must also satisfy Shannon's sampling theorem to capture the dominant spectral information of the field. Thus, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_b < \frac{1}{2\rho_b\nu_{co}}. 
\end{equation}
where $\rho_b$ is an over-sampling parameter to determine the basis function separation and $\nu_{co}$ is the spatial cut-off frequency from the observations. For the Gaussian basis functions, $\phi(\mathbf{r})$, being considered, the basis function width parameter, $\sigma_{\phi}^2$, can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian is another Gaussian given by
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\nu}$, should be chosen to give at least 3 dB attenuation at $\boldsymbol\nu_{co}$ giving
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu}=\frac{2\boldsymbol\nu_c^\top \boldsymbol\nu_c}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_{co}$. Proofs for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.
% Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with at least the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma_{\phi}^2$.
The dynamic field must be spatially oversampled using a larger number of sensors then imposed by frequency analysis considerations, to use frequency analysis to determine the minimum distance and width of the basis functions. Following this, the distance between basis functions is usually larger then the distance between sensors, although this is not a requirement for using this estimation framework. \omg{Ken: can you please write something about what should be done in this case?} For a spatially homogeneous isotropic fields, the basis functions can be placed on a regular grid. Following this, the knowledge of the distance between basis functions directly implies the total number of basis functions required.

@@

	\lambda=\alpha^2( L+\kappa) - L 

@@

where $\alpha$ (determines spread of sigma points), $\beta$ (incorporate prior knowledge of the distribution of $\mathbf{x}$) and $\kappa$ (scaling parameter) are constants (see ~\cite{Haykin2001} for more details). The sigma vectors are propagated through the system equations and weighted to form the predicted mean and covariance. The weights are calculated by 

@@

Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are  used to correct the predicted states. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the predicted cross-covariance matrix of the states is also calculated. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that as the disturbance, $e_t\left(\mathbf{r}\right)$ and the measurement noise $ \boldsymbol{\varepsilon}_t$ are additive, therefore the additive form of the URTS is used, otherwise the state vector should be augmented with the noise terms. The additive form reduces the dimension and the number of sigma points used in the algorithm.

@@

  \begin{figure}
   	\begin{center}
   		\includegraphics[width=0.5\textwidth]{./Graph/MAE.pdf} 
   	\end{center}
   	\caption{The mean of the absolute error over time, gray area shows $\pm$STD.} 
\label{fig:MAE}
   \end{figure}
\begin{figure*}[!th]
\centering 
\subfigure[][]{\includegraphics[width=2in]{./Graph/TrueField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimatedField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimationError90.pdf}}\\
\caption{One snapshot of the spatial field; a: True field; b: Estimated field; c: The absolute error.}
\label{fig:FieldEstimate}
\end{figure*}
